\Title

### 2021-08-25 latex/paper/main.tex
Updated the proof, needs way more editing but it's now complete.

    If at any point in discrete time \(r_n\) exits \(]\underline{r},\overline{r}[\), the orbit 
    in \((q_u,p_u)\) coordinates must have departed \(D\).
    Hence, proving the constrained dynamics gain energy on \(D\) is
    equivalent to finding \(N \in \mathbb{Z}_{> 0}\) such that 
    \(r_{n} \geq \overline{r}\) for all \(n \geq N\).
    Using a perturbation analysis (see Khalil 
    \cite[Theorem 10.1]{khalil_nonlinear}),
    we know there exists some \(I_2 \in \, ]0,I_1]\) and \(\kappa > 0\)
    such that \(\left|R(\theta,r,I)\right| \leq \kappa I^2\) for all 
    \(\theta \in [0,2\pi]\),
    \(r \in \, ]\underline{r},\overline{r}[\),
    and \(I \in [-I_2,I_2]\).
    Taking this fact alongside the assumption that \(S(r_0) \geq \epsilon\),
    we find a lower bound \(r_{n+1} \geq r_n + I\epsilon - \kappa I^2\).
    Choosing \(\gamma > 0\) and \(I^\star \in \, ]0,I_2]\) small enough makes 
    \(I\epsilon - \kappa I^2 \geq \gamma\) for all \(I \in \, ]0,I^\star]\).
    Therefore, \(r_{n} \geq r_0 + n\gamma\), 
    so all solutions of \eqref{eqn:r-discrete} eventually flow past
    \(\overline{r}\).
    We conclude that, for \(I \in \, ]0, I^\star]\), the constrained dynamics gain
    energy on \(D\).

@@

Recall the discrete-time pseudo-radius dynamics \eqref{eqn:r-discrete} from the
proof of Lemma \ref{lemma:energy-regulation-D}. 
To prove the origin is a repeller, we need simply
prove the linearization of \eqref{eqn:r-discrete} is greater than \(1\) at 
\(r_0 = 0\). 
This linearization yields
\[
    P^\prime(0) = 1 + I K Q^\prime(0) + R^\prime(2\pi,0,I)
    ,
\]
where the prime now denotes differentiation with respect to \(r\).
One can show that
%\begin{equation*}
%    \lim \limits_{r_0 \to 0^+}
%    \pdiff{a}{r}(\theta,r_0) 
%    = -\frac{\sqrt{2}}{2} \left(11\sin(\theta)^2 - 6\right)
%    ,
%\end{equation*}
%which means
\begin{align*}
    Q^\prime(0) = \int \limits_0^\pi 
    \lim \limits_{r_0 \to 0^+}\pdiff{a}{r}(\sigma,r_0) d\sigma
    = \frac{\pi}{2\sqrt{2}} > 0
    .
\end{align*}
Since \(R(2\pi,0,I)\) is \(O(I^2)\), so is
\(R^\prime(2\pi,0,I)\).
Hence, it can be written in the form
\(I^2 \tilde{R}(I)\) where \(\tilde{R}(I)\) is smooth and zero at \(I = 0\).
Thus, there exists \(I_0 > 0\) such that 
\(I K Q^\prime(0) + I^2 \tilde{R}(I) > 0\) for all \(I \in \, ]0,I_0]\),
which implies \(P^\prime(0) > 1\).
The origin is therefore a repeller, which means there exists some
(unknown) \(\mu > 0\) where \(]0,\mu]\) is negatively
invariant for \eqref{eqn:r-discrete}, and all solutions in this
interval flow through \(r = \mu\).
Hence, all orbits in \(\mathcal{O}_1\) near the
origin will gain energy up to \(E(\mu,0)\).

@@

Using this fact, we now prove that almost every orbit in \(\mathcal{O}_1\) will
cross the line \(|q_u| = \pi\).
Let \(x(0) \in \mathcal{O}_1 \backslah \Pi^+\) and suppose that the orbit
\(x([0,\infty[)\) does not hit the line \(|q_u| = \pi\).
This orbit must remain in the compact set \(\Sone \times [-K,K]\) for some 
\(K > 0\).
Indeed, suppose otherwise.
Then for arbitrary \(K > 0\), there exists some \(t_K > 0\) such that
\(|p_u(t_K)| = K\), and there exists a \(t_{K/2}\) so that
\(|p_u(t)| >= K/2\) for \(t \in [t_K, t_K + t_{K/2}]\).
Looking at the constrained dynamics \eqref{eqn:acrobot-constrained-dynamics} in
this timespan, we see that \(\dot{p}_u \geq -3mgl\) and
\(\dot{q}_u \geq c_1 p_u + c_2 \geq c_1 K/2 + c_2\) for some constants
\(c_1, c_2 > 0\).
Picking \(K\) large enough guarantees that \(q_u\) covers a distance of \(2\pi\)
over this time interval, and so the orbit hits the line \(|q_u| = \pi\)
(thus providing a contradiction).

Thus, the orbit \(x(t)\) remains within the compact set \(\Sone \times [-K,K]\).
By the Birkhoff theorem, \(x(t)\) approaches a positive limit set, which is
either a closed orbit or the equilibrium \((\pi,0)\).

TODO: closed orbit case

We conclude that almost all orbits beginning in \(\mathcal{O}_1\) will, in
finite time, escape the closure of \(\mathcal{O}_1\) and begin rotating by
crossing the vertical line \(|q_u| = \pi\).

>>>>

    If at any point in discrete time \(r_n\) exits \(]\underline{r},\overline{r}[\), the orbit 
    in \((q_u,p_u)\) coordinates must have departed \(D\).
    Hence, proving the constrained dynamics gain energy on \(D\) is
    equivalent to finding \(N \in \mathbb{Z}_{> 0}\) such that 
    \(r_{n} \geq \overline{r}\) for all \(n \geq N\).
    Using a perturbation analysis (see Khalil 
    \cite[Theorem 10.1]{khalil_nonlinear}),
    we know there exists some \(I_2 \in \, ]0,I_1]\) and \(\kappa > 0\)
    such that \(\left|R(\theta,r,I)\right| \leq \kappa I^2\) for all 
    \(\theta \in [0,2\pi]\),
    \(r \in \, ]\underline{r},\overline{r}[\),
    and \(I \in [-I_2,I_2]\).
    Taking this fact alongside the assumption that \(S(r_0) \geq \epsilon\),
    we find a lower bound \(r_{n+1} \geq r_n + I\epsilon - \kappa I^2\).
    Let \(0 < I^\star < \min\{I_2, \epsilon/\kappa\}\).
    Then for each \(I \in ]0,I^\star]\), the value 
    \(\gamma := I\epsilon - \kappa I^2\) is strictly positive, so
    \(P(r_0) \geq r_0 + \gamma\).
    Therefore, \(r_{n} \geq r_0 + n\gamma\), 
    and all solutions of \eqref{eqn:r-discrete} eventually flow past
    \(\overline{r}\).
    We conclude that, for \(I \in \, ]0, I^\star]\), the constrained dynamics gain
    energy on \(D\).

@@

Recall the discrete-time pseudo-radius dynamics \eqref{eqn:r-discrete} from the
proof of Lemma \ref{lemma:energy-regulation-D}. 
To prove the origin is a repeller, we need simply
prove the linearization of \eqref{eqn:r-discrete} is greater than \(1\) at 
\(r_0 = 0\). 
This linearization yields
\[
    P^\prime(0) = 1 + I K Q^\prime(0) + R^\prime(2\pi,0,I)
    ,
\]
where the prime now denotes differentiation with respect to \(r\).
One can show that
%\begin{equation*}
%    \lim \limits_{r_0 \to 0^+}
%    \pdiff{a}{r}(\theta,r_0) 
%    = -\frac{\sqrt{2}}{2} \left(11\sin(\theta)^2 - 6\right)
%    ,
%\end{equation*}
%which means
\begin{align*}
    Q^\prime(0) = \int \limits_0^\pi 
    \lim \limits_{r_0 \to 0^+}\pdiff{a}{r}(\sigma,r_0) d\sigma
    = \frac{\pi}{2\sqrt{2}} > 0
    .
\end{align*}
Since \(R(2\pi,0,I)\) is \(O(I^2)\), so is
\(R^\prime(2\pi,0,I)\).
Hence, it can be written in the form
\(I^2 \tilde{R}(I)\) where \(\tilde{R}(I)\) is smooth and zero at \(I = 0\).
Thus, there exists \(I_0 > 0\) such that 
\(I K Q^\prime(0) + I^2 \tilde{R}(I) > 0\) for all \(I \in \, ]0,I_0]\),
which implies \(P^\prime(0) > 1\).
The origin is therefore a repeller, which means there exists some
(unknown) \(\mu > 0\) where \(]0,\mu]\) is negatively
invariant for \eqref{eqn:r-discrete}, and all solutions in this
interval flow through \(r = \mu\).
Hence, all orbits in \(\mathcal{O}_1\) near the
origin will gain energy up to (and including) \(E(\mu,0)\).

@@

Using this fact, we now prove that almost every orbit in \(\mathcal{O}_1\) will
cross the line \(|q_u| = \pi\).
Let \(x(0) \in \mathcal{O}_1\) and suppose that the orbit
\(x([0,\infty[)\) does not hit the line \(|q_u| = \pi\).
We claim that this orbit must remain in the compact set 
\(\Sone \times [-K,K]\), for some \(K > 0\).
By contradiction, suppose otherwise.
Then for arbitrary \(K > 0\), there exists some \(t_K > 0\) such that
\(|p_u(t_K)| = K\). 
Looking at the constrained dynamics \eqref{eqn:acrobot-constrained-dynamics},
we see that \(\dot{p}_u \geq -3mgl\) and
\(\dot{q}_u \geq c_1 p_u + c_2\) for some constants \(c_1,c_2 > 0\).
This implies there exists a \(t_{1/2} > 0\) (independent of \(K\)) 
where
\(|p_u(t)| >= K/2\) for \(t \in [t_K, t_K + t_{1/2}]\),
and \(\dot{q}_u \geq c_1 K/2 + c_2\) on this interval.
Picking \(K\) large enough guarantees \(q_u(t)\) covers a distance of \(2\pi\)
over this time interval, and so the orbit must hit the line \(|q_u| = \pi\).
Contradiction.

Thus, the orbit must \(x(t)\) remain within the compact set 
\(\Sone \times [-K,K]\).
By the Birkhoff theorem, \(x(t)\) approaches a positive limit set \(L^+\), 
which is either an invariant closed orbit which crosse the \(q_u\)-axis, 
or the set \(\{(\pi,0)\}\).
By contradiction, suppose \(L^+\) is a closed orbit which 
hits the \(q_u\)-axis at \(\bar{q} < \pi\).
Since \(]0,\mu]\) is negatively invariant for \eqref{eqn:r-discrete}, 
we must have \(\bar{q} \in ]\mu,\pi[\).
The Poincar\'{e} map \eqref{eqn:poincare-map} is smooth on \(]0,\pi[\), and its
image contains \(\bar{q}\), so \(P^{-1}(\bar{q}) \neq \emptyset\).
Take \(q^\prime \in P^{-1}(\bar{q})\). 
If \(q^\prime \in ]0,\mu]\), then \(q^\prime < \bar{q}\).
If \(q^\prime > \mu\), then we know 
from the proof of Lemma \ref{lemma:energy-regulation-D} that
\(P(q^\prime) = \bar{q} \geq q^\prime + \gamma\) for some \(\gamma > 0\), 
so \(q^\prime \leq \bar{q} - \gamma < \bar{q}\).
Since \(q^\prime < \bar{q}\), the orbit starting at \(q^\prime\) stays within
\(\mathcal{O}_1\) and intersects the closed orbit at \(\bar{q}\).
This violates uniqueness of solutions for the constrained dynamics,
so \(L^+\) must be \(\{(\pi,0)\}\), which in turn implies that 
\(x(0) \in \Pi^+\).

We conclude that almost all orbits beginning in \(\mathcal{O}_1\) will
escape the closure of \(\mathcal{O}_1\) and, in finite time, 
begin rotating by crossing the vertical line \(|q_u| = \pi\).
